Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS

### 2011-09-08 ms.tex
B_min proceedure revised, simulation-model adopted to Fermi constraints, text changes propagated

\subsection{Simulation-based approach}
We adopt the universal spectral and spatial gamma-ray model developed by \citet{article:PinzkePfrommer:2010} to estimate the emission from decaying neutral pions that dominates over the IC emission from primary and secondary electrons above 100 MeV in clusters. Given a density profile as, e.g., inferred by cosmological simulations or X-ray observations, the analytic approach models the CR distribution and the associated radiative emission processes from radio to the gamma-ray band. This formalism was derived from high resolution simulations of clusters of galaxies that included radiative hydrodynamics, star formation, supernova feedback, and followed CR physics by tracing the most important injection and loss processes self-consistently while accounting for the CR pressure in the equation of motion \citep{article:PfrommerSpringelEnsslinJubelgas, article:EnsslinPfrommerSpringelJubelgas:2007, article:JubelgasSpringelEnsslinPfrommer:2008}. The results are in line with earlier numerical results on some of the overall characteristics of the CR distribution and the associated radiative emission processes \citep{article:DolagEnsslin, article:MiniatiRyuKangJones:2001, article:Miniati:2003, article:Pfrommer_etal:2007, article:PfrommerEnsslinSpringel:2008, article:Pfrommer:2008}.

@@

We adopt the electron density profile for the Coma cluster that has been inferred from ROSAT X-ray observations \citep{article:BrielHenryBohringer:1992} and model the spatial and spectral distribution of CRs according to the model by \citet{article:PinzkePfrommer:2010} that neglects the contribution of cluster galaxies. Figure \ref{fig:spectrum} shows the expected spectrum of Coma, normalized to the total flux within the virial radius. In our further analysis, we use the gamma-ray flux limits of {\em Fermi} in the energy interval of 1--3 GeV which are most constraining as well as the gamma-ray flux limits of VERITAS in the energy range above 220 GeV.

Figure \ref{fig:xcr} shows the relative CR pressure, $X_{\CR} = \expval{P_{\CR}}/\expval{P_\rmn{th}}$ as a function of radial distance, $R$, from the Coma cluster center (dashed red lines) and contained within $R$ (dashed blue lines). To compute the CR pressure, we assume a low-momentum cutoff of the CR distribution at $q = 0.8\,m_p c$ as found in the cosmological simulations which reflects the high cooling rates at low CR energies. The relative CR pressure rises toward the outer regions on account of the higher efficiency of CR acceleration at the peripheral strong accretion shocks compared to weak central flow shocks. Adiabatic compression of a mixture of CRs and thermal gas disfavours the CR pressure relative to the thermal pressure on account of the softer equation of state of CRs. The weak increase of $X_{\CR}$ toward the core is due to the comparably fast thermal cooling of gas which diminishes the thermal support relative to that in CRs.

Table \ref{table:constraints} shows the expected gamma-ray fluxes within the regions of interest. For the most constraining regions tested (within an aperture of radius 0.2$^{\circ}$), those are a factor of 7.2 below the inferred upper limits of VERITAS (see Table \ref{table:results} and assuming a spectral index of $\alpha=2.3$ which matches the simulated one at energies $E=200$ GeV). Hence we scale the averaged relative CR pressure of our model by that factor, keep the spatial behavior and obtain an integrated limit of the relative CR pressure of $X_{\CR}<0.091$ within 0.2$^{\circ}$ that translates to a limit within the cluster virial radius of $X_\CR<0.13$ (solid lines of Figure \ref{fig:xcr}).

[Anders: add and discuss Fermi limits on XCR]

\subsection{Simplified analytical model}
To complement the simulation-inspired analysis and to facilitate comparison with earlier analytical work, we repeat the data analysis with a simplified analytical model that assumes a power-law CR spectrum and a constant relative CR pressure, i.e., we adopt the {\em isobaric model of CRs}. To be independent of additional assumptions and in line with earlier work in the literature, we do not impose a low-momentum cutoff on the CR distribution function, i.e., we adopt $q=0$. Since a priori, the CR spectral index is unconstrained, we vary it within plausible ranges of $2.1<\alpha_{\rmn{p}}<2.5$. This spectral range is compatible with the radio spectral index of the giant radio halo of Coma after accounting for the spectral steepening at frequencies $\nu\sim5~\rmn{GHz}$ due to the Sunyaev-Zel'dovich effect \citep{article:PfrommerEnsslin:2004b}\footnote{Assuming a magnetic field of 1 $\mu$G, the CR protons responsible for the GHz radio emitting electrons have an energy of $\sim100$ GeV and
  are $\sim$ 20 times less energetic than those CR protons responsible for 200-GeV gamma-ray emission.}. Table \ref{table:constraints} shows the resulting constraints on $X_\CR$. The VERITAS flux upper limits strongly depend on $\alpha$ due to the comparably large lever arm from GeV energies (that dominate the CR pressure) to TeV-energies. The limits within 0.2$^{\circ}$ are most constraining yielding limits on $X_\CR$ between 0.048 and 0.43 (for $\alpha$ varying between 2.1 and 2.5). The {\em Fermi} limits depend only weakly on $\alpha$ and are most constraining for an aperture of 0.4$^{\circ}$.  The best limit on $X_\CR$ of 0.017 is achieved for $\alpha=2.3$.

%This is more constraining by 50\% in comparison to the model derived from simulations. (in comparison to a power law spectrum with $\alpha_\rmn{TeV} \sim 2.2$ down to a cutoff in the normalized momentum of 0.8)

@@

\section{Constraining the Magnetic Field}
\label{sec:B}
We have seen in the previous section that we can obtain an absolute lower limit on the gamma-ray emission in the hadronic model by assuming high magnetic fields, $B>B_\rmn{CMB}$. Hence, we can turn this around by using our upper limit on the gamma-ray emission (and by extension on the CR pressure) to infer a lower limit on the magnetic field that is needed to explain the observed radio emission which, however, assuming the validity of the hadronic model \citep{article:PfrommerEnsslin:2004a}. Lowering the gamma-ray constraint will tighten (increase) the magnetic field limit. Once this gets in conflict with magnetic field measurements by means of other methods, e.g., Faraday rotation measure (RM)\footnote{Generally, Faraday RM analysis of the magnetic-field strength by, e.g., background sources through clusters are degenerate with the magnetic coherence scale and may be biased by the unknown correlation between magnetic and density fluctuations.}, this would challenge the hadronic model of radio halos. The method that we use to constrain the magnetic field inherits a dependence on the assumed spatial structure that we parametrize by
\begin{equation}
\label{eq:B}
B(r) = B_{0} \,\left(\frac{n_{\rmn{e}}(r)}{n_{\rmn{e}}(0)}\right)^{\alpha_B},
\end{equation}
as suggested by Faraday RM studies and numerical magneto-hydrodynamical (MHD) simulations \citep[][and references therein]{article:Bonafede_etal:2010, article:Bonafede_etal:2011}. In fact, the magnetic field in the Coma cluster is among the best constrained, because its proximity permits RM observations of 
seven radio-sources located at projected distances of 50 to 1500 kpc from the cluster center. The best-fit model obtains $B_{0} = 4.7^{+0.7}_{-0.8}\,\mu$G and $\alpha_{B} = 0.5^{+0.2}_{-0.1}$ \citep{article:Bonafede_etal:2010}. We aim for constraining the central field strength, $B_{0}$, which depends on the magnetic decline, $\alpha_{B}$, that we vary (within reasonable ranges of $\Delta\alpha_{B}=0.2$ as suggested from the Faraday RM studies) to address this dependence. We proceed as follows:

\begin{enumerate}
\item Given a model for the magnetic field, we determine the profile of the relative CR pressure, $X_{\CR}(r)$, so that the hadronically produced CR electrons exactly matches the observed radio halo emission over the entire extent. To this end, we de-project the fit to the surface brightness profile (using an Abel integral equation, see Appendix of \citealt{article:PfrommerEnsslin:2004b}) yielding the radio emissivity,
\begin{equation}
\label{eq:Coma:radio}
j_{\nu} (r) = \frac{S_{0}}{2\pi\, r_{\rmn{c}}}\,
\frac{6\beta - 1}{\left(1 + r^{2}/r_{\rmn{c}}^{2}\right)^{3 \beta}}\,
\mathcal{B}\left(\frac{1}{2}, 3\beta\right)
= j_{\nu,0} \left(1 + r^2/r_{\rmn{c}}^{2}\right)^{-3 \beta},
\end{equation}
where $\mathcal{B}$ denotes the beta function. It is generically true for weak magnetic fields ($B<B_{\rmn{CMB}}$) in the outer parts of the Coma halo that the product $X_{\CR}(r)X_{B}(r)$ has to increase by a factor of about 100 toward the radio halo periphery to account for the observed extent. If we were to adopt a steeper magnetic decline such as $\alpha_{B}=0.5$ which produces a flat $X_{B}(r)$, the relative CR pressure would have to rise accordingly by a factor of 100.

\item Given this realization for $X_{\CR}$, we compute the pion-decay gamma-ray surface brightness profile, integrate the flux within a radius of $(0.13, 0.2, 0.4)$ degree, and scale the CR profile in order to match the corresponding VERITAS/{\em Fermi} upper limits. This scaling factor, $X_{\CR,0}$, depends on the CR spectral index, $\alpha$, (assuming a power-law CR population for simplicity) and the radial decline of the magnetic field, $\alpha_{B}$.

\item In the last step, we solve for $B_{0}$ subject to the boundary condition to match the observed synchrotron profile while fixing the profile of $X_{\CR}(r)$ as determined through the previous two steps. Note that for $B_{0} \gg B_{\rmn{CMB}}$ and a radio spectral index of $\alpha_{\nu}=1$, the solution would be degenerate.
\end{enumerate}

Table~\ref{table:Bmin} shows the results on the lower limit of the central magnetic field that range for the VERITAS limits from $B_{0} = 0.4$ to $1.2\,\mu$G and for the {\em Fermi} limits from $B_{0} = 1$ to $7\,\mu$G. Since these lower limits on $B_0$ are below the values favored by Faraday RM for most of the parameter space spanned by $\alpha_B$ and $\alpha$, the hadronic model is still viable. The only exception is the parameter combination of $\alpha_B=0.7$ and $\alpha=2.5$ which starts to be in (weak) tension with Faraday RM data. This indicates the power of future gamma-ray observations of Coma in constraining parameters of the hadronic model.

In addition there are a few interesting remarks in order. (1) For the VERITAS limits, the hardest CR spectral indices correspond to the tightest limits on $B_0$. This is because for an upper limit of CRs around 1.6 TeV and a CR population with a softer spectral index (larger $\alpha$) there is a comparably larger fraction of CRs at 100 GeV available which inject more radio-emitting electrons. In order to reproduce the observed synchrotron emission, this allows for smaller magnetic fields. The opposite is true for the {\em Fermi} upper limits at 1 GeV which probe CRs around a pivot point of 8 GeV: a softer CR spectral index implies a comparably smaller fraction of CRs at 100 GeV and hence decreases the radio-emitting secondary electrons. Equivalently larger magnetic fields are needed to compensate this in order to match the observed synchrotron flux.  (2) For a steeper magnetic decline (larger $\alpha_{B}$), the CR number density needs to be larger to match the observed radio emission profiles which would yield a higher gamma-ray flux so that the upper limits are more constraining. This implies tighter lower limits for $B_{0}$. However, the largest choice of $\alpha_B=0.7$, which represents a flux-freezing magnetic field, is possibly too steep in the outer parts as it neglects the contributions from turbulent dynamos in these dynamically active regions with vorticity injection at corrugated structure formation shocks \citep{article:PfrommerJones:2011}. (3) Interestingly, in all cases, we use the 0.4$^{\circ}$ limit that is always most constraining for our choice of magnetic field configurations.  This is because we need a substantially increasing relative CR pressure profile to match the observed radio profiles (for a given magnetic realization); hence that CR realization produces a larger flux within 0.4$^{\circ}$ in comparison to the ``isobaric CR model'' assumption ($X_{\CR} = \rmn{const.}$) for which the 0.2$^{\circ}$ upper limits are more constraining; hence the
constraints on $B_{0}$ get better for the larger integration area. Physically, this may be achieved by considering CR transport processes such as streaming which produces an increasing $X_{\CR}$-profile due to the larger available phase space in the outer cluster regions.

As a final word, we note that the $X_{\CR}$ profiles always obey the energy condition, i.e., $P_{\CR} < P_{\mathrm{th}}$ over the entire range of the radio halo emission ($< 1$ Mpc) when assuming a central magnetic field of $B_{0}=3\,\mu$G that is at the lower 2-$\sigma$ interval bound inferred from Faraday RM \citep{article:Bonafede_etal:2010}.\footnote{Note that a central magnetic field of $3\,\mu$G corresponds in the Coma cluster to a relative magnetic energy density of $\tilde{X}_B=0.005$.} Typical values at 1 Mpc range from $X_{\CR} = 0.02$ ($\alpha_{B}=0.3$, $\alpha=2.1$) to $X_{\CR} = 1$ ($\alpha_{B}=0.7$, $\alpha=2.5$).\footnote{See Figure 3 in \citet{article:PfrommerEnsslin:2004a} for the entire parameter range assuming minimum energy conditions and \citet{article:PfrommerEnsslin:2004b}, Figure 7 for a parametrization as adopted in this study.} We caution, however, that the minimum-energy condition is violated at the outer radio halo boundary for the range of minimum
magnetic-field values inferred by this study.

@@

\section{Discussion and Conclusions}
VERITAS observed the Coma cluster of galaxies for a total of 18.6 hours of high-quality live time between March and May in 2008.

We employed various approaches to constrain the CR population and magnetic field distribution that are complementary in their assumptions and hence well suited in assessing underlying Bayesian priors in the models. (1) We use a simplified ``isobaric CR model'' that is characterized by a constant CR-to-thermal pressure fraction and has a power-law momentum spectrum.  While this model is physically not justified a priory, it is simple and widely used in the literature and captures some aspects of more elaborate models such as (2) the simulation-based analytical approach of \citet{article:PinzkePfrommer:2010}. This is a ``first principle approach'' that predicts the CR distribution for a given set of assumptions (namely the physics that was simulated). It is powerful since it only requires the density profile as input due to the approximate universality of the CR distribution (when neglecting CR diffusion and streaming which appears, however, necessary to explain the radio halo bimodality). (3) We finally use a pragmatic approach which models the CR and magnetic distribution such as to reproduce the observed emission profile of the Coma radio halo. While this approach is also physically not justified, it is powerful as it shows what the ``correct'' model has to achieve and can point into the direction of the relevant physics.

Within this pragmatic approach we employ two different directions. First, adopting a high magnetic field everywhere in the cluster ($B\gg B_\rmn{CMB}$) yields the minimum gamma-ray flux in the hadronic model of radio halos which we find to be a factor of 20 (60) below the most constraining flux upper limits of {\em Fermi} (VERITAS). Second, by matching the radio emission profile (i.e., fixing the radial CR profile for a given magnetic field model) and by requiring the pion-decay gamma-ray flux to match the flux upper limits (i.e., fixing the normalization of the CR distribution) we obtain lower limits on the magnetic field distribution under consideration. Our limits of the central magnetic field range from $B_{0} = 0.4$ to $1.2\,\mu$G (for VERITAS gamma-ray flux limits) and from $B_{0} = 1$ to $7\,\mu$G (for {\em Fermi} gamma-ray flux limits). Since these lower limits on $B_0$ are below the values favored by Faraday RM, the hadronic model is still viable.

Using the most constraining VERITAS limits, we can constrain the CR-to-thermal pressure, $X_\CR$, to be below 0.048 and 0.43 (for a CR or gamma-ray spectral index, $\alpha$, varying between 2.1 and 2.5).  The best {\em Fermi} flux limit constrains $X_\CR$ to be below 0.017 (for $\alpha=2.3$), only weakly depending on the assumed CR spectral index.

@@

\begin{figure*}
\begin{center}
\scalebox{1.0}{\plotone{f4.eps}}
\end{center}
\caption{Integral flux upper limits (this work, Table X) compared with simulated integrated spectra of VHE gamma-ray emission from the Coma cluster.}
\label{fig:spectrum}
\end{figure*}

>>>>

\subsection{Simulation-based approach}
We adopt the universal spectral and spatial gamma-ray model developed by \citet{article:PinzkePfrommer:2010} to estimate the emission from decaying neutral pions that dominates over the IC emission from primary and secondary electrons above 100 MeV in clusters. Given a density profile as, e.g., inferred by cosmological simulations or X-ray observations, the analytic approach models the CR distribution and the associated radiative emission processes from radio to the gamma-ray band. This formalism was derived from high-resolution simulations of clusters of galaxies that included radiative hydrodynamics, star formation, supernova feedback, and followed CR physics by tracing the most important injection and loss processes self-consistently while accounting for the CR pressure in the equation of motion \citep{article:PfrommerSpringelEnsslinJubelgas, article:EnsslinPfrommerSpringelJubelgas:2007, article:JubelgasSpringelEnsslinPfrommer:2008}. The results are in line with earlier numerical results on some of the overall characteristics of the CR distribution and the associated radiative emission processes \citep{article:DolagEnsslin, article:MiniatiRyuKangJones:2001, article:Miniati:2003, article:Pfrommer_etal:2007, article:PfrommerEnsslinSpringel:2008, article:Pfrommer:2008}.

@@

We adopt the electron density profile for the Coma cluster that has been inferred from ROSAT X-ray
observations \citep{article:BrielHenryBohringer:1992} and model the spatial and spectral
distribution of CRs according to the model by \citet{article:PinzkePfrommer:2010} that neglects the
contribution of cluster galaxies. Figure \ref{fig:spectrum} shows the expected spectrum of Coma,
normalized to the total flux within the virial radius (dotted lines) and the angular apertures tested in
this study (solid lines). The gamma-ray flux limit of {\em Fermi} in the energy interval of 1--3 GeV
($<0.4^\circ$) is most constraining. However, since the model flux within the finite energy range of
1--3 GeV is smaller by a factor of 1.5, it falls below the flux upper limits and does not yet
constrain the simulation model.  In our further analysis, we use the most constraining {\em Fermi}
flux limits in the energy interval of 1--3 GeV as well as the gamma-ray flux limits of VERITAS in
the energy range above 220 GeV.

Figure \ref{fig:xcr} shows the relative CR pressure, $X_{\CR} =
\expval{P_{\CR}}/\expval{P_\rmn{th}}$ as a function of radial distance, $R$, from the Coma cluster
center (red lines) and contained within $R$ (blue lines).  We normalize the radii by the virial
radius, $R_\rmn{vir}$, that we define as the radius of a sphere enclosing a mean density that is 200
times the critical density of the Universe.  To compute the CR pressure, we assume a low-momentum
cutoff of the CR distribution at $q = 0.8\,m_p c$ as found in the cosmological simulations which
reflects the high cooling rates at low CR energies. The relative CR pressure rises toward the outer
regions on account of the higher efficiency of CR acceleration at the peripheral strong accretion
shocks compared to weak central flow shocks. Adiabatic compression of a mixture of CRs and thermal
gas disfavours the CR pressure relative to the thermal pressure on account of the softer equation of
state of CRs. The weak increase of $X_{\CR}$ toward the core is due to the comparably fast thermal
cooling of gas which diminishes the thermal support relative to that in CRs.

Table \ref{table:constraints} shows the expected gamma-ray fluxes within the regions of interest. In
the case of VERITAS, for the most constraining regions tested (within an aperture of radius
0.2$^{\circ}$), those are a factor of 7.2 below the inferred upper limits of VERITAS (see Table
\ref{table:results} and assuming a spectral index of $\alpha=2.3$ which matches the simulated one at
energies $E_\gamma=200$ GeV). Hence we scale the averaged relative CR pressure of our model by that factor,
keep the spatial behavior and obtain an integrated limit of the relative CR pressure of
$X_{\CR}<0.105$ within 0.2$^{\circ}$ that translates to a limit within the cluster virial radius of
$X_\CR<0.152$ (solid lines of Figure \ref{fig:xcr}). Equivalently, the most constraining {\em Fermi}
upper limit in the energy interval of 1--3 GeV ($<0.4^\circ$) is a factor of 1.1 larger than our
model predictions (assuming $\alpha=2.1$ which is very close to the simulated spectral index for the
energy range of 1--3 GeV). Scaling our integrated relative CR pressure profile yields the
constraint of $X_{\CR}<0.018$ within 0.4$^{\circ}$ that translates to a limit within the cluster
virial radius of $X_\CR<0.024$ (dashed lines of Figure \ref{fig:xcr}).

\subsection{Simplified analytical model}
To complement the simulation-inspired analysis and to facilitate comparison with earlier analytical
work, we repeat the data analysis with a simplified analytical model that assumes a power-law CR
spectrum and a constant relative CR pressure, i.e., we adopt the {\em isobaric model of CRs}. To be
independent of additional assumptions and in line with earlier work in the literature, we do not
impose a low-momentum cutoff on the CR distribution function, i.e., we adopt $q=0$. Since a priori,
the CR spectral index is unconstrained, we vary it within plausible ranges of
$2.1<\alpha_{\rmn{p}}<2.5$. This spectral range is compatible with the radio spectral index of the
giant radio halo of Coma after accounting for the spectral steepening at frequencies
$\nu\sim5~\rmn{GHz}$ due to the Sunyaev-Zel'dovich effect
\citep{article:PfrommerEnsslin:2004b}\footnote{Assuming a magnetic field of 1 $\mu$G, the CR protons
  responsible for the GHz radio emitting electrons have an energy of $\sim100$ GeV and are $\sim$ 20
  times less energetic than those CR protons responsible for 200-GeV gamma-ray emission.}. Table
\ref{table:constraints} shows the resulting constraints on $X_\CR$. The VERITAS flux upper limits
strongly depend on $\alpha$ due to the comparably large lever arm from GeV energies (that dominate
the CR pressure) to TeV-energies. The limits within 0.2$^{\circ}$ are most constraining yielding
limits on $X_\CR$ between 0.048 and 0.43 (for $\alpha$ varying between 2.1 and 2.5), with a
constraint of $X_\CR<0.1$ for $\alpha=2.3$ (close to the spectral index predicted by the simulations
around 220 GeV). This limit is more constraining by 50\% in comparison to the model derived from
simulations, which gives $X_\CR<0.15$. This difference is explained by the concave curvature of the
simulated spectrum which accumulates additional pressure toward GeV energies in comparison to a pure
power-law spectrum of $\alpha=2.3$.

The {\em Fermi} limits depend only weakly on $\alpha$ and are most constraining for an aperture of
0.4$^{\circ}$.  The best limit of $X_\CR< 0.017$ is achieved for $\alpha=2.3$ and only slightly
worse for $\alpha\simeq 2.1$ yielding $X_\CR<0.025$.  This is comparable to the $X_\CR$-constraint
in our simulation-based model of $X_\CR<0.024$ (evaluated within the cluster virial
radius). Naturally, with the Fermi limits we probe the region around GeV-energies that dominate the
CR pressure and we do not expect any differences in comparison to the simulation-based models on account
of the concave curvature of the universal CR spectrum found in the simulations.

@@

\section{Constraining the Magnetic Field}
\label{sec:B}
We have seen in the previous section that we can obtain an absolute lower limit on the gamma-ray emission in the hadronic model by assuming high magnetic fields, $B\gg B_\rmn{CMB}$. Hence, we can turn this around by using our upper limit on the gamma-ray emission (and by extension on the CR pressure) to infer a lower limit on the magnetic field that is needed to explain the observed radio emission which, however, assumes the validity of the hadronic model \citep{article:PfrommerEnsslin:2004a}. Lowering the gamma-ray constraint will tighten (increase) the magnetic field limit. Once this gets in conflict with magnetic field measurements by means of other methods, e.g., Faraday rotation measure (RM)\footnote{Generally, Faraday RM analysis of the magnetic-field strength by, e.g., background sources through clusters are degenerate with the magnetic coherence scale and may be biased by the unknown correlation between magnetic and density fluctuations.}, this would challenge the hadronic model of radio halos. The method that we use to constrain the magnetic field inherits a dependence on the assumed spatial structure that we parametrize by
\begin{equation}
\label{eq:B}
B(r) = B_{0} \,\left(\frac{n_{\rmn{e}}(r)}{n_{\rmn{e}}(0)}\right)^{\alpha_B},
\end{equation}
as suggested by Faraday RM studies and numerical magneto-hydrodynamical (MHD) simulations \citep[][and references therein]{article:Bonafede_etal:2010, article:Bonafede_etal:2011}. In fact, the magnetic field in the Coma cluster is among the best constrained, because its proximity permits RM observations of 
seven radio-sources located at projected distances of 50 to 1500 kpc from the cluster center. The best-fit model obtains $B_{0} = 4.7^{+0.7}_{-0.8}\,\mu$G and $\alpha_{B} = 0.5^{+0.2}_{-0.1}$ \citep{article:Bonafede_etal:2010}. We aim for constraining the central field strength, $B_{0}$, which depends on the magnetic decline, $\alpha_{B}$, that we vary (within reasonable ranges of $\Delta\alpha_{B}=0.2$ as suggested from the Faraday RM studies) to address this dependence. We proceed as follows:

\begin{enumerate}
\item Given a model for the magnetic field with $\alpha_B$ and an initial guess for $B_0$, we
  determine the profile of the relative CR pressure, $X_{\CR}(r)$, so that the hadronically produced
  synchrotron emission exactly matches the observed radio halo emission over the entire extent. To
  this end, we de-project the fit to the surface brightness profile of Eqn.~\ref{beta} (using an Abel integral
  equation, see Appendix of \citealt{article:PfrommerEnsslin:2004b}) yielding the radio emissivity,
  \begin{equation}
    \label{eq:Coma:radio}
    j_{\nu} (r) = \frac{S_{0}}{2\pi\, r_{\rmn{c}}}\,
    \frac{6\beta - 1}{\left(1 + r^{2}/r_{\rmn{c}}^{2}\right)^{3 \beta}}\,
    \mathcal{B}\left(\frac{1}{2}, 3\beta\right)
    = j_{\nu,0} \left(1 + r^2/r_{\rmn{c}}^{2}\right)^{-3 \beta},
  \end{equation}
  where $\mathcal{B}$ denotes the beta function. It is generically true for weak magnetic fields
  ($B<B_{\rmn{CMB}}$) in the outer parts of the Coma halo that the product $X_{\CR}(r)X_{B}(r)$
  (where $X_B$ denotes the magnetic-to-thermal energy density) has to increase by a factor of about
  100 toward the radio halo periphery to account for the observed extent. If we were to adopt a
  steeper magnetic decline such as $\alpha_{B}=0.5$ which produces a flat $X_{B}(r)$, the relative
  CR pressure would have to rise accordingly by a factor of 100.

\item Given this realization for $X_{\CR}$, we compute the pion-decay gamma-ray surface brightness
  profile, integrate the flux within a radius of $(0.13, 0.2, 0.4)$ degree, and scale the CR profile
  in order to match the corresponding VERITAS/{\em Fermi} flux upper limits. This scaling factor,
  $X_{\CR,0}$, depends on the CR spectral index, $\alpha$, (assuming a power-law CR population for
  simplicity), the radial decline of the magnetic field, $\alpha_{B}$, and our initial guess for
  $B_0$.

\item We then solve for $B_{0}$ subject to the boundary condition to match the observed synchrotron
  profile while fixing the profile of $X_{\CR}(r)$ as determined through the previous two
  steps. Note that for $B_{0} \gg B_{\rmn{CMB}}$ and a radio spectral index of $\alpha_{\nu}=1$, the
  solution would be degenerate.

\item Inverse Compton cooling of CR electrons on CMB photons introduces a characteristic scale of
  $B_\rmn{CMB}\simeq 3.2\,\mu$G which imprints as a non-linearity on the synchrotron emissivity as a
  function of magnetic field strength. Hence we have to iterate through the previous steps until our
  solution for the minimum magnetic field $B_0$ is converged.
\end{enumerate}

Table~\ref{table:Bmin} shows the results on the lower limit of the central magnetic field that range
for the VERITAS limits from $B_{0} = 0.5$ to $1.4\,\mu$G and for the {\em Fermi} limits from $B_{0}
= 1.1$ to $4.6\,\mu$G.\footnote{Note that a central magnetic field of $3\,\mu$G corresponds in the
  Coma cluster to a relative magnetic energy density of $X_B=0.005$.} Since these lower limits on
$B_0$ are below the values favored by Faraday RM for the entire parameter space spanned by
$\alpha_B$ and $\alpha$, the hadronic model is perfectly viable.  This indicates the power of future
gamma-ray observations of Coma in constraining parameters of the hadronic model.

In addition there are a few interesting remarks in order. (1) For the VERITAS limits, the hardest CR
spectral indices correspond to the tightest limits on $B_0$. This is because for an upper limit of
CRs around 1.6 TeV and a CR population with a softer spectral index (larger $\alpha$) there is a
comparably larger fraction of CRs at 100 GeV available which inject more radio-emitting
electrons. In order to reproduce the observed synchrotron emission, this allows for smaller magnetic
fields. The opposite is true for the {\em Fermi} upper limits at 1 GeV which probe CRs around a
pivot point of 8 GeV: a softer CR spectral index implies a comparably smaller fraction of CRs at 100
GeV and hence decreases the radio-emitting secondary electrons. Equivalently larger magnetic fields
are needed to compensate for this in order to match the observed synchrotron flux.  (2) For a
steeper magnetic decline (larger $\alpha_{B}$), the CR number density needs to be larger to match
the observed radio emission profiles which would yield a higher gamma-ray flux so that the upper
limits are more constraining. This implies tighter lower limits for $B_{0}$.  (3) Interestingly, in
all cases, we use the 0.4$^{\circ}$ limit that is always most constraining for our choice of
magnetic field configurations.  This is because we need a substantially increasing relative CR
pressure profile to match the observed radio profiles (for a given magnetic realization); hence that
CR realization produces a larger flux within 0.4$^{\circ}$ in comparison to the ``isobaric CR
model'' assumption ($X_{\CR} = \rmn{const.}$) for which the 0.2$^{\circ}$ upper limits are more
constraining in the case of VERITAS; hence the constraints on $B_{0}$ get better for the larger
integration area. Physically, this may be achieved by considering CR transport processes such as
streaming which produces an increasing $X_{\CR}$-profile due to the larger available phase space in
the outer cluster regions.

As a final word, in Table~\ref{table:Bmin} we additionally show the
corresponding values for the relative CR pressure (at the largest emission
radius at 1~Mpc) such that the model reproduces the observed radio surface
brightness profile. They should be interpreted as upper limits since they are
derived from flux upper limits. For the {\em Fermi} upper limits, they range
from 0.1 to 0.4; hence the $X_{\CR}$-profiles always obey the energy condition,
i.e., $P_{\CR} < P_{\mathrm{th}}$ over the entire range of the radio halo
emission ($< 1$ Mpc).\footnote{See Figure 3 in
  \citet{article:PfrommerEnsslin:2004a} for the entire parameter range assuming
  minimum energy conditions and \citet{article:PfrommerEnsslin:2004b}, Figure 7
  for a parametrization as adopted in this study.  We caution, however, that
the minimum-energy condition is violated at the outer radio halo boundary for
the range of minimum magnetic-field values inferred by this study.} The
corresponding values for $X_\CR$ in the cluster center are smaller than 0.01 for
the entire parameter space probed in this study. We conclude that the hadronic
model is not challenged by current Faraday RM data and a perfectly viable
possibility in explaining the Coma radio halo emission.

@@

\section{Discussion and Conclusions}
VERITAS observed the Coma cluster of galaxies for a total of 18.6 hours of high-quality live time between March and May in 2008. [{\em add flux limits}]

We employed various approaches to constrain the CR population and magnetic field distribution that
are complementary in their assumptions and hence well suited in assessing underlying Bayesian priors
in the models. (1) We use a simplified ``isobaric CR model'' that is characterized by a constant
CR-to-thermal pressure fraction and has a power-law momentum spectrum.  While this model is
physically not justified a priory, it is simple and widely used in the literature and captures some
aspects of more elaborate models such as (2) the simulation-based analytical approach of
\citet{article:PinzkePfrommer:2010}. This is a ``first principle approach'' that predicts the CR
distribution spectrally and spatially for a given set of assumptions (namely the physics that was
simulated). It is powerful since it only requires the density profile as input due to the
approximate universality of the CR distribution (when neglecting CR diffusion and streaming which
appears, however, necessary to explain the radio halo bimodality). (3) We finally use a pragmatic
approach which models the CR and magnetic distribution such as to reproduce the observed emission
profile of the Coma radio halo. While this approach is also physically not justified, it is powerful
as it shows what the ``correct'' model has to achieve and can point into the direction of the
relevant physics.

Within this pragmatic approach we employ two different directions. First, adopting a high magnetic
field everywhere in the cluster ($B\gg B_\rmn{CMB}$) yields the minimum gamma-ray flux in the
hadronic model of radio halos which we find to be a factor of 20 (60) below the most constraining
flux upper limits of {\em Fermi} (VERITAS). Second, by matching the radio emission profile (i.e.,
fixing the radial CR profile for a given magnetic field model) and by requiring the pion-decay
gamma-ray flux to match the flux upper limits (i.e., fixing the normalization of the CR
distribution) we obtain lower limits on the magnetic field distribution under consideration. Our
limits of the central magnetic field range from $B_{0} = 0.5$ to $1.4\,\mu$G (for VERITAS 
flux limits) and from $B_{0} = 1.1$ to $4.6\,\mu$G (for {\em Fermi} flux limits). Since
all these lower limits on $B_0$ are below the values favored by Faraday RM, $B_{0} =
4.7^{+0.7}_{-0.8}\,\mu$G \citep{article:Bonafede_etal:2010}, the hadronic model is a very attractive
explanation of the Coma radio halo.

Applying our simplified ``isobaric CR model'' to the most constraining VERITAS limits, we can
constrain the CR-to-thermal pressure, $X_\CR$, to be below 0.048 and 0.43 (for a CR or gamma-ray
spectral index, $\alpha$, varying between 2.1 and 2.5). We obtain a constraint of $X_\CR<0.1$ for
$\alpha=2.3$, the spectral index predicted by simulations at energies around 220 GeV.  This limit is
more constraining by a factor of 1.5 in comparison to the simulation-based model which gives
$X_\CR<0.15$.  This difference is due to the concave curvature of the simulated spectrum which
accumulates additional pressure toward GeV energies in comparison to a pure power-law spectrum of
$\alpha=2.3$.

The {\em Fermi} flux limits constrain $X_\CR$ to be below 0.017--0.025 (for $\alpha$ varying between
2.3 and 2.1), only weakly depending on the assumed CR spectral index. Assuming $\alpha=2.1$ which is
very close to the simulated spectral index for the energy range of 1--3 GeV, we obtain an almost
coincident constraint with our simulation-based model within the virial radius of
$X_\CR<0.024$. That constraint improves to $X_{\CR}<0.018$ for an aperture of 0.4$^\circ$
corresponding to a physical scale of $R \simeq R_{200}/3 \simeq 660$~kpc. These are encouraging
results in that we constrain the CR pressure (of a phase that is fully mixed with the ICM) to be at
most a small fraction of the overall pressure, $<0.025$. As a result, hydrostatic cluster masses and
the total Comptonization parameter due to the Sunyaev-Zel'dovich effect attain at most a very small
bias due to CRs.

@@

\begin{figure*}
\begin{center}
\scalebox{1.0}{\plotone{f4.eps}}
\end{center}
\caption{Integral flux upper limits (this work, Tables~\ref{table:results} and \ref{table:fermi})
  for different integration radii (shown with different colors), assuming $\alpha=2.1$. These are
  compared to simulated integrated spectra of VHE gamma-ray emission, $E_\gamma
  F_\gamma(>E_\gamma)$, from the Coma cluster (again within different apertures). Note that the {\em
    Fermi} limit for the energy interval of 1--3 GeV within the aperture of 0.4$^\circ$ is most
  constraining. Since the model flux within the finite energy range of 1--3 GeV is smaller by a
  factor of 1.5, it falls below the flux upper limits and does not yet constrain the simulation
  model.}
\label{fig:spectrum}
\end{figure*}
